---
title: "Clark-Izzi-ADA-homework-2"
output: html_document
---
Challenge 1:
```{r message=FALSE, include=FALSE, messages=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(sciplot)
library(radiant)
library(mosaic)
```
Filter dataset, add decade column, plot runtimeMinutes histogram for each decade.
```{r}
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)
s <- filter(d, startYear >= 1920 & startYear <= 1979 & runtimeMinutes < 240)
s <- mutate(s, "decade" = paste(substr(startYear,3,3), "0s", sep = ""))
p <- ggplot(s, aes(x=runtimeMinutes), color = factor(decade)) + geom_histogram() + facet_wrap(~decade, ncol = 3) # wrap data by decade into 3 columns
p
```
Population mean and population standard deviation (using sdpop() from {radiant}) in runtimeMinutes for each decade
```{r}
results <- s %>% group_by(decade) %>%
  summarise(
    popMean = mean(runtimeMinutes, na.rm = TRUE),
    popSD = sd(runtimeMinutes, na.rm = TRUE),
  )
results
```
Single sample of 100 movies from each decade: sample mean and standard deviation, plus estimate of SE around population mean
```{r}
results_samp <- tibble(decade = character(), sampMean = numeric(), sampSD = numeric(), estSE = numeric())
for (d in unique(s$decade)){
  samp <- sample_n(filter(s, decade == d), size = 100, replace = FALSE)
  x <- mean(samp$runtimeMinutes, na.rm = TRUE)
  y <- sd(samp$runtimeMinutes, na.rm = TRUE)
  z <- y / sqrt(100) # estimate pop SE from SD of sample
  results_samp <- add_row(results_samp, decade = d, sampMean = x, sampSD = y, estSE = z)
}
results_samp
```
Compare estimates to actual population mean and calculated SE based on population standard deviation. (I added to the population results table from earlier.) Slightly different but comparable to sample stats!
```{r}
results <- mutate(results, "calcSE" = popSD/sqrt(100)) # expected SE for sample size of 100
results
```
Sampling distribution of mean runtimeMinutesfor each decade. Had trouble actually getting them to print from inside the loop - maybe my computer is too slow? Commented it out for now.
```{r eval=FALSE, include=FALSE}
for (d in unique(s$decade)){
  x <- filter(s, decade == d)
  samp_dist_mean <- do(10000) * mean(~runtimeMinutes, data = sample_n(x, size = 100, replace = FALSE))
  #mean_plot <- histogram(~mean, data = samp_dist_mean, xlab = paste("Sampling Distribution for the\nMean of runtimeMinutes: ", d))
  samp_dist_sd <- do(10000) * sd(~runtimeMinutes, data = sample_n(x, size = 100, replace = FALSE))
  #sd_plot <- histogram(~sd, data = samp_dist_sd, xlab = paste("Sampling Distribution for the\nStandard deviation of runTimeMinutes: ", d))
  #plot_grid(mean_plot, sd_plot)
}
```

Challenge 2:

What is the probability that she will hear 13 or fewer calls during any given session? 0.1425978
```{r}
l = 18
ppois(q = 13, lambda = l)
```
What is the probability that she will hear no calls in a session? 1.522998e-08
```{r}
dpois(x = 0, lambda = l)
```
What is the probability that she will hear exactly 7 calls in a session? 0.00185002
```{r}
dpois(x = 7, lambda = l)
```
What is the probability that she will hear more than 20 calls in a session? (1-ppois?) 0.2692798
```{r}
1-ppois(q = 20, lambda = l)
```
Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 40.
```{r}
p <- plotDist(
  "pois",
  lambda = l,
  xlim = c(0, 40),
  main = paste0("Poisson Distribution\nwith lambda=", l), xlab = "x",
  ylab = "Pr(X=x)"
)
p
```
Using the rpois() function, simulate 520 results from this distribution (i.e., 10 full years of Saturday monitoring sessions).
```{r}
pois_dist <- rpois(n = 520, lambda = l)
```
Plot the simulated results. How do your simulated results compare to the shape of the probability mass function you plotted above? 520 may not be a large enough sample (running a few times generated variable results), but overall similar to the probability mass function with a peak density around 18.
```{r}
histogram(pois_dist,
  xlim = c(0,40)
)
```
Challenge 3:
```{r}
f2 <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/zombies.csv"
d <- read_csv(f2, col_names = TRUE)
```
Population mean and standard deviation for each variable:
```{r}
height_m <- mean(d$height)
height_sd <- sdpop(d$height)
weight_m <- mean(d$weight)
weight_sd <- sdpop(d$weight)
age_m <- mean(d$age)
age_sd <- sdpop(d$age)
zombies_m <- mean(d$zombies_killed)
zombies_sd <- sdpop(d$zombies_killed)
education_m <- mean(d$years_of_education)
education_sd <- mean(d$years_of_education)
```
Scatterplots of height and weight in relation to age. Both increase as age increases, with an especially clear positive relationship between height and age.
```{r}
p <- ggplot(data = d, aes(
  x = age,
  y = height,
))
p <- p + geom_point(na.rm = TRUE)
p
p2 <- ggplot(data = d, aes(
  x = age,
  y = weight,
))
p2 <- p2 + geom_point(na.rm = TRUE)
p2
```
Histograms and Q-Q plots for each quantitative variable:
Height, weight, and age are normally distributed.
Distributions of zombies killed and years of education are slightly skewed (more people kill fewer zombies and have fewer years of education).
```{r}
histogram(d$height, center = height_m, main = "Height")
qqnorm(d$height, main = "Height")
qqline(d$height, col = "gray")

histogram(d$weight, center = weight_m, main = "Weight")
qqnorm(d$weight, main = "Weight")
qqline(d$weight, col = "gray")

histogram(d$age, center = age_m, main = "Age")
qqnorm(d$age, main = "Age")
qqline(d$age, col = "gray")

histogram(d$zombies_killed, center = zombies_m, main = "Zombies Killed")
qqnorm(d$zombies_killed, main = "Zombies Killed")
qqline(d$zombies_killed, col = "gray")

histogram(d$years_of_education, center = education_m, main = "Years of Education")
qqnorm(d$years_of_education, main = "Years of Education")
qqline(d$years_of_education, col = "gray")
```
Sample 1 subset of 30 survivors.
```{r}
s <- sample_n(d, size = 30, replace = FALSE)
s <- select(s, height, weight, age, zombies_killed, years_of_education)
samp30 <- tibble(variable = character(), mean = numeric(), sd = numeric(), se = numeric(), ci_lower = numeric(), ci_upper = numeric())
for (i in seq_along(s)) {
  m <- mean(s[[i]])
  sd <- sd(s[[i]])
  se <- sd / sqrt(30)
  ci_L <- m - qnorm(1 - 0.05 / 2) * se
  ci_U <- m + qnorm(1 - 0.05 / 2) * se
  samp30 <- add_row(samp30, variable = (colnames(s)[i]), mean = m, sd = sd, se = se, ci_lower = ci_L, ci_upper = ci_U)
}
samp30
```
Sample another 99 samples of 30 survivors. Means and standard errors of samping distrubtions are similar to corresponding values generated above! Weight, height, and age look sort of normally distributed, while zombies killed and years of education still look slightly non-normal.
```{r}
set.seed(1)
height_dist <- do(99) * mean(~height, data = sample_n(d, size = 30, replace = FALSE))
weight_dist <- do(99) * mean(~weight, data = sample_n(d, size = 30, replace = FALSE))
age_dist <- do(99) * mean(~age, data = sample_n(d, size = 30, replace = FALSE))
zombies_dist <- do(99) * mean(~zombies_killed, data = sample_n(d, size = 30, replace = FALSE))
education_dist <- do(99) * mean(~years_of_education, data = sample_n(d, size = 30, replace = FALSE))
mean(~mean, data = height_dist)
sd(~mean, data = height_dist)
histogram(~mean, data = height_dist, main = "Height")
mean(~mean, data = weight_dist)
sd(~mean, data = weight_dist)
histogram(~mean, data = weight_dist, main = "Weight")
mean(~mean, data = age_dist)
sd(~mean, data = age_dist)
histogram(~mean, data = age_dist, main = "Age")
mean(~mean, data = zombies_dist)
sd(~mean, data = zombies_dist)
histogram(~mean, data = zombies_dist, main = "Zombies Killed")
mean(~mean, data = education_dist)
sd(~mean, data = education_dist)
histogram(~mean, data = education_dist, main = "Years of Education")

```


